### 2019-11-24 ic-journal/ccs.tex
Cumulative fixes based on discussions on three questions

\section{CCS}
\label{ss:ccs}

Here we recall the core theory of CCS.
We assume a possibly infinite set of names $\mathscr{L} = \{a, b,
\ldots\}$ yielding input and $\overline{\mbox{output}}$ labels (actions), plus a special invisible
action $\tau \notin \mathscr{L}$, and another possibly infinite
set of agent variables $A, B, \ldots \in \mathscr{X}$.
The class of \hl{all} CCS \hl{terms (or expressions)} is then inductively
defined from $\nil$ (the terminated process) and agent variables by the operators
of \emph{prefixing} (.), \emph{parallel composition} ($|$),
\emph{summation}
(or \emph{binary choice}, $+$),
\emph{restriction} ($\nu$), \emph{relabeling} ($[\cdot]$) and
\emph{recursion} ($\texttt{rec}$):
\begin{equation*}
\begin{array}{cccl}
\mu  & := & \tau\!\!\!\! & \midd \; a  \; \midd \;  \outC a  \\
P  & := & \nil\!\!\!\! & \midd \;  \mu . P \; \midd \;  P_1 |  P_2 \; \midd  \;
P_1 + P_2 \; \midd 
(\res L)\, P
\; \midd \; P\; [\mathit{rf}]  % relabelling
\;  \midd \;  A \; \midd \; \recu A  P
\end{array}
\end{equation*}
In our presentation of CCS, the restriction operator takes a set of labels $L
  \subseteq \mathscr{L}$, rather than a single one.  
The relabeling operator takes a relabeling function
  $\mathit{rf} \colon \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\} \rightarrow \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\}$, that can  handle multiple actions including $\tau$. A valid
    relabeling function $\mathit{rf}$ must however 
satisfy $\mathit{rf}(\tau) =
    \tau$ and $\forall l\in\mathscr{L} \cup \overline{\mathscr{L}}.\, \mathit{rf}(\overline{l}) =
    \overline{\mathit{rf}(l)}$ (with $\bar{\bar l} = l$ for all $l \in
    \mathscr{L}$).
We sometimes omit a trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
A CCS process $P$ may evolve to another one, say $P'$, under
  an action $\mu$, written by $P \arr{\mu} P'$.
The transition semantics of CCS processes is given by means of
a Labeled Transition System (LTS) expressed in Structural Operational
Semantics (SOS) rules shown in Fig.~\ref{f:LTSCCS}.
(The symmetric versions of the SOS rules for
parallel composition and binary choice are omitted.)
A CCS processes uses only \emph{guarded sums} if all occurrences of
sums are in the form $a.P + b.Q$. 
The \emph{immediate derivatives} of a
process $P$ are elements of the set $\{P' \st P \arr\mu P' \mbox{for some $\mu$}\}$.
We use $\ell$ to range over visible actions (i.e.~inputs or outputs, excluding  $\tau$)
and $\alpha, \mu$  to range over all actions.
\begin{figure*}[t]
\begin{center}
\vskip .1cm
 $\displaystyle{  \over  \mu.  P \arr\mu
P } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P + Q   \arr\mu
P'  } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P | Q   \arr\mu
P' | Q } $  $ \hb$   
\hskip .3cm
  $\; \;$  $\displaystyle{ P \arr{ a}P' \hk \hk  Q
\arr{\outC a }Q'  \over  P|  Q \arr{ \tau} P'
|  Q'  }$ 
\\
\vspace{.2cm}
$\displaystyle{ P \arr{\mu}P' \over
 (\res L\!)\, P \arr{\mu} (\res L\!)\, P'} $ $ \mu, \outC\mu \not\in L$
$ \hb$
%
$\displaystyle{ P \sub {\recu A P} A \arr{\mu}P' \over
 \recu A P   \arr{ \mu} P'  } $
\hskip .5cm  
$\displaystyle{ P \arr{\mu} P' \over
 P \;[r\!f] \arr{r\!f(\mu)} P' \;[r\!f]} $
$ \hb$ %  &
\end{center}
\caption{Structural Operational Semantics of CCS}
\label{f:LTSCCS}
\end{figure*}

@@

\paragraph{Further notations}
We let $\R$, $\S$ range over binary relations, sometimes using 
 infix notation for them; e.g., 
$P \,\R\, Q$ means $(P,Q) \in \R$.
We use a tilde, as in  $\til P$, to denote tuples of processes with
countably many elements (thus the tuple may also be infinite.)
All notations are extended to tuples componentwise,
e.g., $\til P \,\R\, \til Q$ means $P_i \,\R\, Q_i$ for each  
index $i$ of the tuples $\til P$ and $\til Q$.
We use the symbol $\DSdefi$ for abbreviations. For instance, $P \DSdefi G $, where
$G$ is some expression, means that $P$ stands for the expression $G$.
If $\leq$ is a preorder, then  $\geq$  is its inverse (and
conversely).

>>>>

\section{Calculus of Communicating Systems (CCS)}
\label{ss:ccs}

Here we recall the core theory of CCS.
We assume a possibly infinite set of names $\mathscr{L} = \{a, b,
\ldots\}$ yielding input and $\overline{\mbox{output}}$ labels (actions), plus a special invisible
action $\tau \notin \mathscr{L}$, and another possibly infinite
set of agent variables \hl{$\mathscr{X} = \{X, Y, \ldots\}$}.
The class of \hl{all} CCS \hl{terms (or expressions)} is then inductively
defined from $\nil$ (the terminated process) and agent variables by the operators
of \emph{prefixing} (.), \emph{parallel composition} ($|$),
\emph{summation}
(or \emph{binary choice}, $+$),
\emph{restriction} ($\nu$), \emph{relabeling} ($[\cdot]$) and
\emph{recursion} ($\texttt{rec}$):
\begin{equation*}
\begin{array}{cccl}
\mu  & := & \tau\!\!\!\! & \midd \; a  \; \midd \;  \outC a  \\
P  & := & \nil\!\!\!\! & \midd \;  \mu . P \; \midd \;  P_1 |  P_2 \; \midd  \;
P_1 + P_2 \; \midd 
(\res L)\, P
\; \midd \; P\; [\mathit{rf}]  % relabelling
\;  \midd \;  X \; \midd \; \recu X P
\end{array}
\end{equation*}
In our presentation of CCS, the restriction operator takes a set of labels $L
  \subseteq \mathscr{L}$, rather than a single one.  
The relabeling operator takes a relabeling function
  $\mathit{rf} \colon \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\} \rightarrow \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\}$, that can  handle multiple actions including $\tau$. A valid
    relabeling function $\mathit{rf}$ must however 
satisfy $\mathit{rf}(\tau) =
    \tau$ and $\forall l\in\mathscr{L} \cup \overline{\mathscr{L}}.\, \mathit{rf}(\overline{l}) =
    \overline{\mathit{rf}(l)}$ (with $\bar{\bar l} = l$ for all $l \in
    \mathscr{L}$).
We sometimes omit a trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
A CCS process $P$ may evolve to another one, say $P'$, under
  an action $\mu$, written by $P \arr{\mu} P'$.
The transition semantics of CCS processes is given by means of
a Labeled Transition System (LTS) expressed in Structural Operational
Semantics (SOS) rules shown in Fig.~\ref{f:LTSCCS}.
(The symmetric versions of the SOS rules for
parallel composition and binary choice are omitted.)
A CCS processes uses only \emph{guarded sums} if all occurrences of
sums are in the form $a.P + b.Q$. 
The \emph{immediate derivatives} of a
process $P$ are elements of the set $\{P' \st P \arr\mu P' \mbox{for some $\mu$}\}$.
We use $\ell$ to range over visible actions (i.e.~inputs or outputs, excluding  $\tau$)
and $\alpha, \mu$  to range over all actions.
\begin{figure*}[t]
\begin{center}
\vskip .1cm
 $\displaystyle{  \over  \mu.  P \arr\mu
P } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P + Q   \arr\mu
P'  } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P | Q   \arr\mu
P' | Q } $  $ \hb$   
\hskip .3cm
  $\; \;$  $\displaystyle{ P \arr{ a}P' \hk \hk  Q
\arr{\outC a }Q'  \over  P|  Q \arr{ \tau} P'
|  Q'  }$ 
\\
\vspace{.2cm}
$\displaystyle{ P \arr{\mu}P' \over
 (\res L\!)\, P \arr{\mu} (\res L\!)\, P'} $ $ \mu, \outC\mu \not\in L$
$ \hb$
%
$\displaystyle{ P \sub {\recu X P} X \arr{\mu}P' \over
 \recu X P \arr{\mu} P'} $
\hskip .5cm  
$\displaystyle{ P \arr{\mu} P' \over
 P \;[r\!f] \arr{r\!f(\mu)} P' \;[r\!f]} $
$ \hb$ %  &
\end{center}
\caption{Structural Operational Semantics of CCS}
\label{f:LTSCCS}
\end{figure*}

@@

\paragraph{Further notations}
We let $\R$, $\S$ range over binary relations, sometimes using 
 infix notation for them, \hl{e.g.} $P \,\R\, Q$ means $(P,Q) \in \R$.
We use a tilde, as in  $\til P$, to denote tuples of processes with
countably many \hl{elements.}
All \hl{relation} notations can be extended to tuples componentwise,
e.g., $\til P \,\R\, \til Q$ means $P_i \,\R\, Q_i$ for each index $i$ of the tuples $\til P$ and $\til Q$.
We use $\DSdefi$ to \hl{define} abbreviations. For instance, $P \DSdefi G $, where
$G$ is some expression, means that $P$ stands for the expression $G$.
If $\leq$ is a preorder, then  $\geq$  is its inverse (and
conversely).

### 2019-11-24 ic-journal/expansion.tex
Cumulative fixes based on discussions on three questions

\emph{Strong} bisimulation and strong bisimilarity ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in the above
definition with normal transitions $Q \arr \mu Q'$ (similar for the other case).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.\  $P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hl{do} not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, if % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}

The above definition also brings up a proof technique for proving rooted
bisimilarity from a given bisimulation. The next
lemma plays an
important role in proving some key results in this paper:
\begin{lemma}[Rooted bisimilarity by bisimulation]
\label{l:obsCongrByWeakBisim}
Given a (weak) bisimulation $\RR$, suppose two processes $P$ and $Q$
satisfy the following properties:
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \RR Q'$.
\end{enumerate}
Then $P$ and $Q$ are rooted bisimilar, i.e.~$P \approx^c Q$.
\end{lemma}

One basic property of rooted bisimiarity is congruence,
indeed the coarsest congruence contained in bisimilarity
(c.f. Section~\ref{s:coarsest} and \hl{\mbox{Section~\ref{ss:context}}} for more details):
\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS, and it is the
coarsest congruence contained in $\wb$.
\end{theorem}

>>>>

\emph{Strong} bisimulation and strong bisimilarity ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in the above
definition with normal transitions $Q \arr \mu Q'$ (similar for the other case).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.~$P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hl{do} not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, if % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}
% This new sentence maintains a link with Definition 2.1
\hl{It can be proved that, $P_1 \rapprox Q_1 \wedge P_2 \rapprox Q_2 \Longrightarrow
P_1 + P_2 \rapprox Q_1 + Q_2$.}
The above definition also brings up a proof technique for proving rooted
bisimilarity from a given bisimulation. The next
lemma plays an
important role in proving some key results in this paper:
\begin{lemma}[Rooted bisimilarity by bisimulation]
\label{l:obsCongrByWeakBisim}
Given a (weak) bisimulation $\RR$, suppose two processes $P$ and $Q$
satisfy the following properties:
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \RR Q'$.
\end{enumerate}
Then $P$ and $Q$ are rooted bisimilar, i.e.~$P \approx^c Q$.
\end{lemma}

One basic property of rooted bisimiarity is congruence,
indeed the coarsest congruence contained in bisimilarity
(See Section~\ref{s:coarsest} and \hl{\mbox{Section~\ref{ss:context}}} for more details):
\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS, and it is the
coarsest congruence contained in $\wb$.
\end{theorem}

### 2019-11-24 ic-journal/formal.tex
Cumulative fixes with highlights
Cumulative fixes based on discussions on three questions

\subsection{Higher Order Logic}

@@

The terms of the HOL logic are expressions that denote elements of the
sets denoted by types. There're four kinds of terms in the HOL
logic. There can be described approximately by the BNF grammar in
Fig. \ref{fig:hol-terms}.

@@

The deductive system of the HOL is specified by eight primitive
derivative rules: (c.f.~\cite{hollogic} for more details.)
\begin{enumerate}
\item Assumption introduction (\texttt{ASSUME});
\item Reflexivity (\texttt{REFL});
\item $\beta$-conversion (\texttt{BETA\_CONV});
\item Substitution (\texttt{SUBST});
\item Abstraction (\texttt{ABS});
\item Type instantiation (\texttt{INST\_TYPE});
\item Discharging an assumption (\texttt{DISCH});
\item Modus Ponens (\texttt{MP}).
\end{enumerate}
All proofs are eventually reduced to applications of the above rules,
which also give the semantics of two foundamental
logical connectives, the equality ($=$) and implication
($\Rightarrow$). The rest logical connectives and first-order
quantifiers, including the logical true (\HOLinline{\HOLConst{T}}) and false (\HOLinline{\HOLConst{F}}), are
further defined as $\lambda$-functions: (one rarely needs to care
things at this level, however.)
\begin{equation*}
\begin{array}{l}
\turn \T       =  ((\lquant{x_{\ty{bool}}}x) =
               (\lquant{x_{\ty{bool}}}x))    \\
\turn \forall  =  \lquant{P_{\alpha\fun\ty{bool}}} P =
                    (\lquant{x}\T ) \\
\turn \exists  =  \lquant{P_{\alpha\fun\ty{bool}}} P({\hilbert}\, P) \\
\turn \F  =  \uquant{b_{\ty{bool}}} b  \\
\turn \neg    =  \lquant{b} b \imp \F \\
\turn {\wedge}  =  \lquant{b_1\ b_2}\uquant{b} (b_1\imp (b_2 \imp b)) \imp b \\
\turn {\vee}  =  \lquant{b_1\ b_2}\uquant{b} (b_1 \imp b)\imp ((b_2 \imp b) \imp b) \\
\turn \OneOne  =  \lquant{f_{\alpha \fun\beta}}\uquant{x_1\ x_2}
                    (f\ x_1 = f\ x_2)  \imp (x_1 = x_2) \\
\turn \Onto  =  \lquant{f_{\alpha\fun\beta}}
                  \uquant{y}\equant{x} y = f\ x \\
\turn \TyDef  =  \lambda P_{\alpha\fun\ty{bool}}\
                  rep_{\beta\fun\alpha}.\;
                  \OneOne\ rep \ \wedge{}\  (\uquant{x}P x = (\equant{y} x = rep\ y))
\end{array}
\end{equation*}
It deserves to mention that, the last logical constant above,
$\TyDef$, can be used to define new HOL types as bijections of
subsets of existing types. (c.f.~\cite{Melham:1989dk} for more details.)
HOL's \texttt{Datatype} package~\cite{Melham:1991, holdesc} automates
this tedious process, and can be used to define needed types in CCS.

@@

Notice that, usually the above four axioms are the only axioms allowed
in conventional formalisation projects in HOL: adding new axioms may
either break the logic \hl{consistency} or make the whole
\hl{formalisation} work less
convincible. \hl{In general,} the theorem prover guarantees that,
without adding new axioms, if both the definitions and the statements
of theorems \hl{were} correct, then the \hl{proofs of these theorems} must
be \hl{also} correct.
% This last sentence doesn't have a good link with the previous context,
% disabled. --Chun
%
%% This is also the case here: SOS rules of CCS are
%% \emph{not} axioms but consequences of the inductive relation\hl{, the
%% CCS transition $\rightarrow$}.

>>>>

\subsection{Higher Order Logic (HOL)}

@@

The terms of the HOL logic are expressions \hl{which} denote elements of the
sets denoted by types. There're four kinds of terms in the HOL
logic. \hl{They} can be described (in simplified forms) by the BNF grammar in
Fig.~\ref{fig:hol-terms}. (See~\ref{hollogic} for a complete description.)

@@

The deductive system of the HOL is specified by eight primitive
derivative rules: (See~\cite{hollogic} for more details.)
\begin{enumerate}
\item Assumption introduction (\texttt{ASSUME});
\item Reflexivity (\texttt{REFL});
\item $\beta$-conversion (\texttt{BETA\_CONV});
\item Substitution (\texttt{SUBST});
\item Abstraction (\texttt{ABS});
\item Type instantiation (\texttt{INST\_TYPE});
\item Discharging an assumption (\texttt{DISCH});
\item Modus Ponens (\texttt{MP}).
\end{enumerate}
All proofs are eventually reduced to applications of the above rules,
which also give the semantics of two foundamental
logical connectives, the equality ($=$) and implication
($\Rightarrow$). The rest logical connectives and first-order
quantifiers, including the logical true (\HOLinline{\HOLConst{T}}) and false (\HOLinline{\HOLConst{F}}), are
further defined as $\lambda$-functions: (one rarely needs to care
things at this level, however.)
\begin{equation*}
\begin{array}{l}
\turn \T       =  ((\lquant{x_{\ty{bool}}}x) =
               (\lquant{x_{\ty{bool}}}x))    \\
\turn \forall  =  \lquant{P_{\alpha\fun\ty{bool}}} P =
                    (\lquant{x}\T ) \\
\turn \exists  =  \lquant{P_{\alpha\fun\ty{bool}}} P({\hilbert}\, P) \\
\turn \F  =  \uquant{b_{\ty{bool}}} b  \\
\turn \neg    =  \lquant{b} b \imp \F \\
\turn {\wedge}  =  \lquant{b_1\ b_2}\uquant{b} (b_1\imp (b_2 \imp b)) \imp b \\
\turn {\vee}  =  \lquant{b_1\ b_2}\uquant{b} (b_1 \imp b)\imp ((b_2 \imp b) \imp b) \\
\turn \OneOne  =  \lquant{f_{\alpha \fun\beta}}\uquant{x_1\ x_2}
                    (f\ x_1 = f\ x_2)  \imp (x_1 = x_2) \\
\turn \Onto  =  \lquant{f_{\alpha\fun\beta}}
                  \uquant{y}\equant{x} y = f\ x \\
\turn \TyDef  =  \lambda P_{\alpha\fun\ty{bool}}\
                  rep_{\beta\fun\alpha}.\;
                  \OneOne\ rep \ \wedge{}\  (\uquant{x}P x = (\equant{y} x = rep\ y))
\end{array}
\end{equation*}
It deserves to mention that, the last logical constant above,
$\TyDef$, can be used to define new HOL types as bijections of
subsets of existing types~\cite{Melham:1989dk}.
HOL's \texttt{Datatype} package~\cite{Melham:1991, holdesc} automates
this tedious process, and can be used for \hl{defining types needed for} CCS.

@@

Notice that, usually the above four axioms are the only axioms allowed
in conventional formalisation projects in HOL: adding new axioms
\hl{manually}, may break the logic \hl{consistency}.
\hl{In general,} the theorem prover guarantees that,
without adding new axioms, if both the definitions and the statements
of theorems are correct, then the \hl{proofs of these theorems} must
be \hl{also} correct.
% This last sentence doesn't have a good link with the previous context,
% disabled. --Chun
%
%% This is also the case here: SOS rules of CCS are
%% \emph{not} axioms but consequences of the inductive relation\hl{, the
%% CCS transition $\rightarrow$}.

### 2019-11-24 ic-journal/sos.tex
Cumulative fixes based on discussions on three questions

The CCS formalisation starts with type definitions for action,
relabeling and the CCS term itself. We use the type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' (\texttt{'b} or
$\beta$ is a \emph{type variable}) for all visible actions. The
cardinality and the exact elements depend on the type variable $\beta$
supplied by the end user.
All actions are divided into input and $\overline{\text{output}}$ ones.
This type is defined by HOL's \texttt{Datatype} package in the
following syntax:
\begin{lstlisting}
Datatype: Label = name 'b | coname 'b
End
\end{lstlisting}
For instance, if the type $\beta$ is instantiated by the type
\texttt{string}, then HOL terms \HOLinline{\HOLConst{name} \HOLStringLit{a}} and
\HOLinline{\HOLConst{coname} \HOLStringLit{b}} denote the input action $a$ and output
action \hl{$\overline{b}$}, respectively.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Action}}'' is the
union of all visible actions, plus the invisible action $\tau$
(\texttt{tau}). For instace, the input action $a$ and output
action \hl{$\overline{b}$} of type \hl{\mbox{``\HOLinline{\HOLTyOp{string} \HOLTyOp{Action}}''}}
are denoted by \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{name} \HOLStringLit{a}\ensuremath{)}} and \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{coname} \HOLStringLit{b}\ensuremath{)}}, respectively.

The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'', accounting for the CCS
expressions, where \texttt{'a} (or $\alpha$) is the type
variable\footnote{The order \hl{between} type variables $\alpha$ and $\beta$
    \hl{can be} arbitrary. Our choice is aligned with the literature.
For instance, $\mathrm{CCS}(h,k)$ is the CCS subcalculus that can use at most $h$ constants
and $k$ actions~\cite{gorrieri2017ccs}. To formalize theorems on
such $\mathrm{CCS}(h,k)$, the required CCS type can be retrieved by instantiating the type
variables $\alpha$ and $\beta$ in ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'' with types
having the corresponding cardinalities $h$ and $k$.} for
agent variables, is then defined inductively: (``\HOLinline{\ensuremath{\beta} \HOLTyOp{Relabeling}}'' is the type of all relabeling
  functions. \hl{Relabeling is formalised but is not to be} discussed in this paper.)
\begin{lstlisting}
Datatype: CCS = nil
              | var 'a
              | prefix ('b Action) CCS
              | sum CCS CCS
              | par CCS CCS
              | restr (('b Label) set) CCS
              | relab CCS ('b Relabeling)
              | rec 'a CCS
End
\end{lstlisting}
The above definitions allow us to use terms like \texttt{nil} and
\texttt{sum P Q} in HOL. Their correspondences with conventional CCS notations
in the literature, are given in Table~\ref{tab:ccsoperator}, where
most CCS operators have also the more readable abbreviation forms,
either for end user or for \TeX{} outputs. (All
  formal theorems and definitions in this paper
  are generated from HOL4. Also, by default, all theorems are fully
  specialised, removing outermost universal quantifiers.)

>>>>

The CCS formalisation starts with type definitions for action,
relabeling and the CCS term itself. We use the type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}''
for all labels (i.e. visible actions), where the type variable $\beta$ corresponds
to $\mathscr{L}$ (the set of names for labels) used at the beginning
of Section~\ref{ss:ccs}. The
cardinality and the exact elements of this type depend on the type variable $\beta$
instantiated by the end user \hl{or further formalisations built upon the
present one.}
All \hl{labels} are divided into input and $\overline{\text{output}}$ ones.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' is defined by HOL's \texttt{Datatype} package in the
following syntax:
\begin{lstlisting}
Datatype: Label = name 'b | coname 'b
End
\end{lstlisting}
For instance, if the type $\beta$ is instantiated by \texttt{string}
(the type of all strings), then HOL terms \HOLinline{\HOLConst{name} \HOLStringLit{a}} and
\HOLinline{\HOLConst{coname} \HOLStringLit{b}} denote the input action $a$ and output
action \hl{$\overline{b}$}, respectively.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Action}}'' is the
union of all visible actions, plus the invisible action $\tau$
(\texttt{tau}). For instace, the input action $a$ and output
action \hl{$\overline{b}$} of type \hl{\mbox{``\HOLinline{\HOLTyOp{string} \HOLTyOp{Action}}''}}
are denoted by \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{name} \HOLStringLit{a}\ensuremath{)}} and \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{coname} \HOLStringLit{b}\ensuremath{)}}, respectively.

The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'', accounting for the CCS
expressions, has two type variables \hl{$\alpha$ and $\beta$ corresponding%
\footnote{The order \hl{between} type variables $\alpha$ and $\beta$
    \hl{can be} arbitrary. Our choice is aligned with the literature.
For instance, $\mathrm{CCS}(h,k)$ is the CCS subcalculus that can use at most $h$ constants
and $k$ actions~\cite{gorrieri2017ccs}. To formalize theorems on
such $\mathrm{CCS}(h,k)$, the required CCS type can be retrieved by instantiating the type
variables $\alpha$ and $\beta$ in ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'' with types
having the corresponding cardinalities $h$ and $k$.}
to $\mathscr{X}$ (the set of agent variables) and $\mathscr{L}$ (the
set of names used for labels), respectively, used at the beginning of} Section~\ref{ss:ccs}.
%
The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'' \hl{is defined inductively by HOL's \texttt{Datatype} package:}
(Here ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Relabeling}}'' is the type of all relabeling
  functions. \hl{Relabeling is formalised but is not to be} discussed in this paper.)
\begin{lstlisting}
Datatype: CCS = nil
              | var 'a
              | prefix ('b Action) CCS
              | sum CCS CCS
              | par CCS CCS
              | restr (('b Label) set) CCS
              | relab CCS ('b Relabeling)
              | rec 'a CCS
End
\end{lstlisting}
The above definitions allow us to use terms like \texttt{nil} and
\texttt{sum P Q} in HOL. Their correspondences with conventional CCS notations
in the literature, are given in Table~\ref{tab:ccsoperator}, where
most CCS operators have also the more readable abbreviation forms,
either for end user or for \TeX{} outputs. (All
  formal theorems and definitions in this paper
  are generated from HOL4. Also, by default, all theorems are fully
  specialised, removing outermost universal quantifiers.)
